---
title: "Reproducable Research Project 1"
author: "Sean Broderick"
date: "September 26, 2016"
output: html_document
---

Download the Data here

```{r setup, include=TRUE, echo=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"C:/Users/brodo/Documents/R/data.zip")
unzip(zipfile="C:/Users/brodo/Documents/R/data.zip", exdir = "C:/Users/brodo/Documents/R")
```


Import the data set, update the date variable to date,plot steps by day, then get the mean and median steps per day

```{r include=TRUE,echo=TRUE}
setwd("C:/Users/brodo/Documents/R")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
DailySteps <- aggregate(activity$steps,by=list(Year=activity$date),FUN=sum)


plot1 <- function() {
hist(DailySteps$x,, main="Daily Steps Taken", col="green", xlab="Daily Steps Taken")
dev.copy(png, file="plot1.png", width=480, height=480)
dev.off()
}
plot1()

DailyClean <- DailySteps[!(is.na(DailySteps$x)),]
mean(DailyClean$x)
median(DailyClean$x)
```


some data cleaning is done here before the average step for each 5 minute interval 



```{r include=TRUE, echo=TRUE}
CleanActivity <- activity[!(is.na(activity$steps)),]
AverageInterval <- aggregate(CleanActivity$steps,by=list(Interval=CleanActivity$interval),FUN=mean)

plot2 <- function() {
plot(AverageInterval$Interval, AverageInterval$x, xlab="5 Minute Intervals", ylab="Steps Taken", type="l", col="brown")
dev.copy(png, file="plot2.png", width=480, height=480)
dev.off()
}
plot2()
```


A simple average was used to impute the missing step records.The plot follows with all records populated.


```{r, include=TRUE,echo=TRUE}
AverageInterval[which.max(AverageInterval$x),]
library(mice)
md.pattern(activity)
library(VIM)
missing<- activity[(is.na(DailySteps$x)),]
non <- na.omit(activity)
missing[is.na(missing$steps), "steps"] <- mean(na.omit(activity$steps))
ActivityNoMissing <- rbind(missing,non)
DailyStepsNoMissing <- aggregate(ActivityNoMissing$steps,by=list(Date=ActivityNoMissing$date),FUN=sum)
plot3 <- function() {
hist(DailyStepsNoMissing$x, main="Daily Steps Taken", col="magenta", xlab="Daily Steps Taken")
dev.copy(png, file="plot3.png", width=480, height=480)
dev.off()
}
plot3()
```


Here you can find some data manipulation and the plot of the average 5 minute intervals weekday vs weekend
```{r,include=TRUE, echo=TRUE}
non$day <- weekdays(non$date)
weekend <- non[non$day %in% c("Saturday","Sunday"),]
weekday <- non[non$day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]


plot4 <- function() {
par(mfrow = c(2,1))
weekdayinterval <- aggregate(weekday$steps,by=list(Interval=weekday$interval),FUN=mean)
weekendinterval <- aggregate(weekend$steps,by=list(Interval=weekend$interval),FUN=mean)
plot(weekdayinterval$Interval, weekdayinterval$x,main="Average Weekday Steps", col="blue", xlab = '', ylab = '', type = "l")
plot(weekendinterval$Interval, weekendinterval$x,main="Average Weekend Steps", col="green", xlab = '', ylab = '', type = "l")
dev.copy(png, file="plot4.png", width=480, height=480) 
dev.off()
}
plot4()
```