---
title: "Reproducable Research Project"
author: "Sean Broderick"
date: "September 24, 2016"
output: html_document
---
Here, we are going to download data and import into R; a user will not to point to their computer with the download.file, unzip and setwd commands.

```{r}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
setwd("C:/Users/brodo/Documents/R")
download.file(url,"C:/Users/brodo/Documents/R/data.zip")
unzip(zipfile="C:/Users/brodo/Documents/R/data.zip", exdir = "C:/Users/brodo/Documents/R")
activity <- read.csv("activity.csv")
```

Here, I am making a histogram showing the Total Number of Steps taken Each Day. Also included will be the median and mean steps taken per day.
```{r}
setwd("C:/Users/brodo/Documents/R")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
DailySteps <- aggregate(activity$steps,by=list(Year=activity$date),FUN=sum)

plot1 <- function() {
hist(DailySteps$x,, main="Daily Steps Taken", col="green", xlab="Daily Steps Taken")
dev.copy(png, file="plot1.png", width=480, height=480)
dev.off()
}
 plot1()

DailyClean <- DailySteps[!(is.na(DailySteps$x)),]
mean(DailyClean$x)
median(DailyClean$x)

```

Here I am doing some data manipulation and then plotting the average steps for 5 minute intervals throughtout the designated time frame
```{r}
CleanActivity <- activity[!(is.na(activity$steps)),]
AverageInterval <- aggregate(CleanActivity$steps,by=list(Interval=CleanActivity$interval),FUN=mean)

plot2 <- function() {
plot(AverageInterval$Interval, AverageInterval$x, xlab="5 Minute Intervals", ylab="Steps Taken", type="l", col="brown")
dev.copy(png, file="plot2.png", width=480, height=480)
dev.off()
}
plot2()
```
Here,  I impute the missing step intervals by placing the mean steps from all non missing observations into the missing step interval.
```{r}
AverageInterval[which.max(AverageInterval$x),]
missing<- activity[(is.na(DailySteps$x)),]
non <- na.omit(activity)
missing[is.na(missing$steps), "steps"] <- mean(na.omit(activity$steps))
ActivityNoMissing <- rbind(missing,non)
DailyStepsNoMissing <- aggregate(ActivityNoMissing$steps,by=list(Date=ActivityNoMissing$date),FUN=sum)
plot3 <- function() {
hist(DailyStepsNoMissing$x, main="Daily Steps Taken", col="magenta", xlab="Daily Steps Taken")
dev.copy(png, file="plot3.png", width=480, height=480)
dev.off()
}
plot3()
```
The code below plots the average steps taken in the 5 minute intervals comparing weekdays from weekends
```{r}
non$day <- weekdays(non$date)
weekend <- non[non$day %in% c("Saturday","Sunday"),]
weekday <- non[non$day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]


plot4 <- function() {
par(mfrow = c(2,1))
weekdayinterval <- aggregate(weekday$steps,by=list(Interval=weekday$interval),FUN=mean)
weekendinterval <- aggregate(weekend$steps,by=list(Interval=weekend$interval),FUN=mean)
plot(weekdayinterval$Interval, weekdayinterval$x,main="Average Weekday Steps", col="blue", xlab = '', ylab = '', type = "l")
plot(weekendinterval$Interval, weekendinterval$x,main="Average Weekend Steps", col="green", xlab = '', ylab = '', type = "l")
dev.copy(png, file="plot4.png", width=480, height=480) 
dev.off()
}
plot4()

```

